Characterization of gaps in directly bonded Si compound optics

### 2014-04-15 AO_bondedSi.tex
incorporate DTJ edits up to page 9, overhaul table 1

The University of Texas diffractive optics group produces diffraction gratings on silicon using microlithographic techniques \cite{1998SPIE.3354..201J,2010SPIE.7739E.146W}.  These gratings are used either in transmission (i.e. as grisms \cite{2008SPIE.7014E..77D, 2010SPIE.7739E.123G}), or in reflection (i.e. as immersion gratings, \cite{2007ApOpt..46.3400M, 2010SPIE.7739E.146W,2012SPIE.8450E..2SG}.  These gratings operate throughout the infrared, from 1.2 to 38 $\mu$m, and find application in ground-based (IGRINS \cite{2010SPIE.7735E..54Y, 2012SPIE.8450E..2SG}), airborne (SOFIA FORCAST \cite{2008SPIE.7014E..77D}), and space (JWST NIRCAM \cite{2005SPIE.5904...21B,2010SPIE.7739E.123G}) applications.  To date, our group has produced the gratings on large (up to 100 mm diameter and 30 mm thick) monolithic pieces of monocrystalline silicon and we shape the substrates by sawing, grinding, and polishing to form the final devices \cite{2010SPIE.7739E.146W}.  The next generation of astronomy and earth science spectrographs, however, will need optical devices larger and more precise than what can be made with existing technology.  The proposed Giant Magellan Telescope \cite{2012SPIE.8444E..1HJ} Near Infrared Spectrograph \cite{2006SPIE.6269E.143J,2010SPIE.7735E..87L} (GMTNIRS) will need immersion gratings with thicknesses up to 60 mm and lengths up to 200 mm.  Substrates of the required size cannot fit into conventional processing equipment like reactive ion etchers and electron-beam lithography tools.  Large, thick Si substrates are also prohibitively expensive to experiment on.  Silicon direct bonding \cite{1986JAP....60.2987S,2012SPIE.8450E..2TV} offers advantages by allowing us to produce the diffraction grating on a thin (0.5$-$10 mm) low-mass Silicon puck that can later be bonded to a fine polished Si prism substrate.  Other groups producing Si diffractive optics are also experimenting with this approach (see \cite{2012SPIE.8450E..2TV} and \url{www.lumella.com}).

@@

Finite gaps will degrade the performance of a final device, either through optical aberrations or transmission loss.  The effect of the gap size on the IR transmission is the topic of this paper.  The effect of the gap on optical aberrations is beyond the scope of this paper.  A gap might also degrade the mechanical integrity of a final compound optic.  In practice, the part must survive thermal shocks and shake tests.  The mechanical properties of the gap are also beyond the scope of this paper.

There are two main solutions for reducing risk attributable to gaps.  We will refer to the two approaches as the \emph{strength approach} and the \emph{resilience approach}.  The \emph{strength approach} is to prevent bubbles from forming in the first place, specifically by removing all possible contaminants from the environment before bonding.  Typically this approach involves disciplined handling protocols, and rigorous multi-stage cleaning steps.  The reason the strength approach is so hard is that hydrocarbons serve as catalysts for gap formation\cite{1991JaJAP..30..615M}.  Hydrocarbons are common in Si wafer handling equipment, so avoiding hydrocarbons is challenging.  Unconventional micro-cleanroom approaches have been effective at meeting this goal \cite{1989JaJAP..28L2141L}.  The \emph{resilience approach} is simply to accept the uncomfortable likelihood that bubbles will occur, and then do something to remove them.  The most cited way to get rid of gaps is to anneal them\cite{Mitani1990}.  We describe the current understanding of both approaches in Section \ref{secHistory}.

Whatever technique one uses to remove or reuce the effects of bubbles, it is essential to have a quantitative method by which to assess accurately the optical impact of the gaps before and after treatment.  This paper introduces a technique for directly measuring gaps at silicon-silicon interfaces.  Other techniques for measuring and mapping gaps have already been described elsewhere \cite{1992JEMat..21..669M}.  One key technique is IR Newton rings from IR imaging\cite{Mitani1990}.  This paper gives deeper understanding of the infrared interference technique seen in IR images as Newton rings.  The number and morphology of Newton rings yields information about the lateral and axial extent of gaps. Specifically, dark-light fringes manifest when the gap is comparable to the vacuum wavelength or larger.  Our technique is IR interference in the limit of small gaps.

There are other techniques for mapping gaps.  X-ray topography \cite{1992JEMat..21..669M,1994JaJAP..33....6H} exploits the change in optical density along lines of sight with gaps to allow precise and high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but provides only coarse estimates of the axial extent.  Ultrasound microscopy exploits the density change at the gap interface to produce reflections of sound waves and map the lateral gap dimensions but produces limited information about the axial dimension \cite{2000RScI...71.1869G}.

Here we describe an exploitation of the IR interference technique with new precision instrumentation in the near infrared to measure the axial extent of gaps at the few nanometer size scale.  In this paper we show that gaps as small as 35 nm cause an unacceptable degradation in transmission performance.  Because the size scales probed are so small, they are comparable to the range over which molecular (van der Waals) forces are acting.  It is conceivable that the strength of the bond energy can be constrained from the \emph{in situ} measurement of the separation and that therefore the optical technique can provide not only a measure of the transmission but also an indication of the mechanical robustness of the bond.  We do not pursue further the question of mechanical robustness.  By measuring both the spatial and axial extent of any gaps, the new metrology technique informs revisions of the bonding strategy that can improve both the optical and mechanical performance.

@@

The foundation paper on Si bonding came in 1986\cite{1986JAP....60.2987S}.  After that, there was a sequence of papers using IR imaging to look for evidence of defects in the bond.  \cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G} directly monitored the bond propagation with IR imaging.  \cite{1988JaJAP..27L2364S, 1989JaJAP..28L2141L} demonstrated gap-free bonding.  \cite{1989JaJAP..28L2141L} directly measured the bond surface energy.  \cite{Mitani1990} discussed the causes and remedies for bubble formation.  Specifically, the IR bubble density changes as a function of anneal temperature\cite{1992JEMat..21..669M}, with bubbles disappearing at temperatures greater than 1100$^\circ$ C.  The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy\cite{feijoo1994}, which lent credibility to the idea of distinct temperature phases and evidence for Si$-$H bonds\cite{1995ApPhA..61..101R}.  A robust picture of the chemistry as a function of anneal temperature was ultimately fleshed out\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.  Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation\cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, and moisture and defects \cite{2001JAP....89.6013L}.  Notably, \cite{JJAP.37.4197} found that surface roughness above 1.3 nm RMS results in poor bonding quality in their sample of wafers etched with Argon.  Results from studies of surface topography are inconclusive\cite{2001JOptA...3...85G}.  Nowhere in the literature did we find a technique similar to the one presented here, probably because the technique relies on the high precision of relatively new IR capable spectroscopy that wasn't widely available until this decade.

@@

So there will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles.  For every obviously detectable IR bubble, there are thousands more particles that could unacceptably degrade the transmission performance.  One key idea that saves the day, is the intuition that the areal fill factor scales at least as fast as $D^2$.  Imagine that the particles serve as miniature tentpoles.  The volume beneath the tentpole looks like a cone.  Since the number density goes down as roughly $D^{-2}$, (per the equation above) the total area covered by particles might be constant in $log(D)$.  Simply put, the big particles cover just as much area as the small particles, even though there are many more small particles.  It is further conceivable that the areal fill factor attributable to a tent pole particle scales even faster than $D^2$ since surface energy probably scales inversely with distance of the pieces (\`a la VanDerWaals).

@@

\subsection{Surface roughness or large-scale aberrations}
The second pitfall to consider is that the stiff non-flat Si surfaces will not conform upon contact leaving the complement of the aberrations as an air gap. This issue is much less of a problem for wafers compared to our unconventional thick ($\sim$ 30 mm) Si applications.  Similarly, surface roughness will also limit the bonding strength  \cite{JJAP.37.4197,2001JOptA...3...85G}.  The details of this aspect of the bonding process are outside the scope of this paper.  In short, most of our thick substrates are optically polished to much less than $\lambda = 632$ nm peak to valley over the 3 to 4 inch diameter clear aperture.  We expect such well-polished substrates to have no deleterious impact on the bonding process.  Some experimental substrates had larger surface error, of order a few waves of surface error.  These substrates were reserved for process development, since we anticipated that bonding might progress either not at all, or with large interface gaps.

@@

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.95\columnwidth]{figs/VG03patternSchematic.pdf}}
\caption{Schematic of part VG03 \label{figVG03pattern}.  The four inch diameter VG03 is comprised of two 3.3 mm thick silicon pucks.  We patterned a one inch diameter hole on one of the pucks.  The depth was 4.0 $\mu$m as indicated by Dektak profilometry.  We also patterned a low-fill factor petal pattern surrounding the hole.  The purpose for these patterns was to have gaps of known dimensions against which we could test the optical metrology technique we describe in this paper.  The technique can reveal the axial extent and fill factor of an embedded gap.}
\end{figure}

We evaluated our metrology method by imbedding gaps of known sizes in the interface.  Here we describe the process of how we made the gaps and how we characterized their axial extent.  The three important characteristics of the gaps are their axial extent, their lateral areal extent, and their areal fill factor.  We bored divots in silicon with inductively coupled plasma etching.  The tool was a Unaxis.  We used two different gases with low and high etch rates.  CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 micron/minute.  The area exposed to the process gases was about 1 inch diameter circles, or 15 mm $\times$ 20 mm rectangles.  We protected the unexposed areas with thick photoresist.  We achieved etch depths of 15$\pm$5 nm and 4000$\pm$100 nm, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.  Figure \ref{figX} shows a schematic of VG03 and its pattern.  In a second round of experiments we achieved gap sizes of 14$-$95 nm on four substrates.  For measurements taken with the Veeco, uncertainties were constructed by inspecting the histogram of topology, as shown in Figure \ref{fig}.  The large-scale distortions were removed from the topology by masking and flattening to ensure that the uncertainties accurately reflect the distribution of measured heights.  The stylus profiliometry measured values were simply assigned an uncertainty of 5\%, which was consistent with our intuition and experience.  Table \ref{tabbondexper} lists part numbers and descriptions for our bonded substrates.

@@

\begin{table}[h!]
\caption{UTexas Si bonding experiments \label{tabbondexper}}
\begin{center}
    \begin{tabular}{ c c c c c c c}
    \hline
    Name & Bonded/DSP & Diameter & Thickness & Anneal Temp & Date & Pattern notes \\ 
    - & - & mm & mm & $^\circ$C & MM/DD/YYY & - \\ 
        \hline
    VG01 & Bonded  & 100  &$1$ $\times$ 2 & 23  & 2/13/2013 &  \\
    VG02 & Bonded & 100 & $3.3$ $\times$ 2 &  23  & 2/13/2013 &  \\
    VG03 & Bonded & 100 & $3.3$ $\times$ 2 &  23 & 2/14/2013 & 4 $\mu$m gap 100\% fill factor hole \& petal pattern\\    
    VG04 & Bonded & 100 & $1$ $\times$ 2 &  23 & 2/14/2013 & 15 nm gap, $\sim50\%$ fill factor\\        
    VG05 & DSP & 100 & 1&  23 & 2/15/2013 & cleaved  \\
    VG06 & DSP & 100 & 3.5 &  23 & 2/15/2013 & \\
    VG07 & - & - & - &  - & - & \\
    VG08 & - & - & - &  - & - & \\
    VG09 & Bonded & 75 & 1.5? & 23 & 6/21/2013 & mesh C, 49 $\pm$6 nm, 50\% fill factor\\
    VG15 & Bonded & 100 & 1.5? &  23 & 6/21/2013 & mesh C, M, F, 14 $\pm$2 nm, 50\% fill factor\\
    VG17 & Bonded & 75 & 6.0? &  23 & 6/21/2013 & mesh C, 64 $\pm$6 nm, 50\% fill factor\\
    VS20 & Bonded & 100 & 1.5? &  23 & 6/21/2013 & mesh C, M, F, 95 $\pm$5 nm, 50\% fill factor\\
    \hline
    \end{tabular}
\end{center}
\end{table}

\section{Substrate preparation}
It is important to clean the surfaces before bonding to minimize the particle density.  The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler.  The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.  We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.  The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.  There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.  We measured the large scale surface flatness of the pucks.  The Fisba2 interferometer has a 2 inch diameter beam at $\lambda=$632.8 nm.  We found a typical peak to valley surface flatness of 2 waves over the central 2 inch diameter.

>>>>

The University of Texas diffractive optics group produces diffraction gratings on silicon using microlithographic techniques \cite{1998SPIE.3354..201J,2010SPIE.7739E.146W}.  These gratings are used either in transmission (i.e. as grisms \cite{2008SPIE.7014E..77D, 2010SPIE.7739E.123G}), or in reflection (i.e. as immersion gratings, \cite{2007ApOpt..46.3400M, 2010SPIE.7739E.146W,2012SPIE.8450E..2SG}.  These gratings operate throughout the infrared, from 1.2 to 38 $\mu$m, and find application in ground-based (IGRINS \cite{2010SPIE.7735E..54Y, 2012SPIE.8450E..2SG}), airborne (SOFIA FORCAST \cite{2008SPIE.7014E..77D}), and space (JWST NIRCAM \cite{2005SPIE.5904...21B,2010SPIE.7739E.123G}) applications.  To date, our group has produced the gratings on large (up to 100 mm diameter and 30 mm thick) monolithic pieces of monocrystalline silicon and we shape the substrates by sawing, grinding, and polishing to form the final devices \cite{2010SPIE.7739E.146W}.  The next generation of astronomy and earth science spectrographs, however, will need optical devices larger and more precise than what can be made with existing technology.  The proposed Giant Magellan Telescope \cite{2012SPIE.8444E..1HJ} Near Infrared Spectrograph \cite{2006SPIE.6269E.143J,2010SPIE.7735E..87L} (GMTNIRS) will need immersion gratings with thicknesses up to 60 mm and lengths up to 200 mm.  Substrates of the required size cannot fit into conventional processing equipment like reactive ion etchers and electron-beam lithography tools.  Large, thick Si substrates are also prohibitively expensive to experiment on.  Silicon direct bonding \cite{1986JAP....60.2987S,2012SPIE.8450E..2TV} offers advantages by allowing us to produce the diffraction grating on a thin (0.5$-$10 mm) low-mass Silicon wafer or puck that can later be bonded to a fine polished Si prism substrate.  Other groups producing Si diffractive optics are also experimenting with or using this approach (see \cite{2012SPIE.8450E..2TV} and \url{www.lumella.com}).

@@

Finite gaps will degrade the performance of a final device, either through optical aberrations or transmission loss.  The effect of the gap size on the IR transmission is the topic of this paper.  The effect of the gap on optical aberrations is beyond the scope of this paper.

A gap might also degrade the mechanical integrity of a final compound optic.  In practice, the part must survive thermal shocks and shake tests.  The gap size scales we can probe with our optical technique are comparable to the range over which molecular (van der Waals) forces are acting.  It is conceivable that the strength of the bond energy can be constrained from the \emph{in situ} measurement of the separation and that therefore our optical technique can provide not only a measure of the transmission but also an indication of the mechanical robustness of the bond.  By measuring both the spatial and axial extent of any gaps, the new metrology technique informs revisions of the bonding strategy that can improve both the optical and mechanical performance.  We do not investigate the mechanical robustness further.

There are two main solutions for reducing risk attributable to gaps.  The first approach is to prevent bubbles from forming in the first place, specifically by removing all possible contaminants from the environment before bonding.  Typically this approach involves disciplined handling protocols, and rigorous multi-stage cleaning steps.  The reason the approach is so hard is that hydrocarbons can serve as catalysts for gap formation\cite{1991JaJAP..30..615M}, and that these hydrocarbons are common in Si wafer handling equipment.  Unconventional micro-cleanroom approaches have been effective at meeting the goal of bubble-free bonding\cite{1989JaJAP..28L2141L}.  The second approach is to accept the likelihood that gaps will occur, and take steps to remove them.  The most cited way to get rid of gaps is to anneal them\cite{Mitani1990}.  We describe the current understanding of both approaches in Section \ref{secHistory}.

Whatever technique one uses to remove or reduce the effects of gaps, it is essential to have a quantitative method by which to assess accurately the optical impact of the gaps before and after treatment.  This paper introduces a technique for directly measuring gaps at silicon-silicon interfaces.  Mitani and G\"oselle review other techniques for measuring and mapping gaps\cite{1992JEMat..21..669M}.  One key technique is IR Newton rings from IR imaging\cite{Mitani1990}.  This paper gives deeper understanding of the infrared interference technique seen in IR images as Newton rings.  The number and morphology of Newton rings yields information about the lateral and axial extent of large gaps. Specifically, dark-light fringes manifest when the gap is comparable to the vacuum wavelength or larger.  Our technique is IR interference in the limit of small gaps down to 20 nm or less.

There are other techniques for mapping gaps.  X-ray topography \cite{1992JEMat..21..669M,1994JaJAP..33....6H} exploits the change in optical density along lines of sight with gaps to allow precise and high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but provides only coarse estimates of the axial extent.  In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but produces limited information about the axial dimension \cite{2000RScI...71.1869G}.

Here we describe an application of the IR interference technique with new precision instrumentation in the near infrared to measure the axial extent of gaps at the few nanometer size scale.  In this paper we show that gaps as small as 35 nm cause an unacceptable degradation in transmission performance.

@@

The foundation paper on Si bonding came in 1986\cite{1986JAP....60.2987S}.  After that, there was a sequence of papers using IR imaging to look for evidence of defects in the bond.  \cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G} directly monitored the bond propagation with IR imaging.  \cite{1988JaJAP..27L2364S, 1989JaJAP..28L2141L} demonstrated gap-free bonding in a micro-cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.  Our group has not attempted the micro-cleanroom technique owing to the challenge of spin-cleaning large (30 mm thick) Si substrates.  It is not clear whether the custom micro-cleanroom technique has gained wide usage.  Notably, the   \cite{1989JaJAP..28L2141L} directly measured the bond surface energy.  \cite{Mitani1990} discussed the causes and remedies for bubble formation.  Specifically, the IR bubble density changes as a function of anneal temperature\cite{1992JEMat..21..669M}, with bubbles disappearing at temperatures greater than 1100$^\circ$ C.  The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy\cite{feijoo1994}, which lent credibility to the idea of distinct temperature phases and evidence for Si$-$H bonds\cite{1995ApPhA..61..101R}.  A robust picture of the chemistry as a function of anneal temperature was ultimately fleshed out\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.  Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation\cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, and moisture and defects \cite{2001JAP....89.6013L}.  Notably, \cite{JJAP.37.4197} found that surface roughness above 1.3 nm RMS results in poor bonding quality in their sample of wafers etched with Argon.  Results from studies of surface topography are inconclusive\cite{2001JOptA...3...85G}.  Nowhere in the literature did we find a technique similar to the one presented here, probably because the technique relies on the high precision of relatively new IR capable spectroscopy that wasn't widely available until this decade.

@@

So there will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles.  For every obviously detectable IR bubble, there are thousands more particles that could unacceptably degrade the transmission performance.  One key idea that saves the day, is the intuition that the areal fill factor scales at least as fast as $D^2$.  Imagine that the particles serve as miniature tentpoles.  The volume beneath the tentpole looks like a cone.  Since the number density goes down as roughly $D^{-2}$, (per the equation above) the total area covered by particles might be constant in $log(D)$.  Simply put, the big particles cover just as much area as the small particles, even though there are many more small particles.  It is further conceivable that the areal fill factor attributable to a tent pole particle scales even faster than $D^2$ since surface energy probably scales inversely with distance of the pieces (\`a la VanDerWaals).  The small particles are even less of a problem than the large particles, since the small ($\sim10-60$ nm) particles will have lower average etalon reflectivity than larger (500-1000 nm) particles.  This last point will become clearer in later sections.

@@

\subsection{Surface roughness or large-scale non-conformance}
Another possible source of gaps at the bond is inherent roughness in the initial surfaces that leads to a failure to conform.  The second pitfall to consider is that the stiff non-flat Si surfaces will not conform upon contact leaving the complement of the aberrations as an air gap. Unconventional thick ($\sim$ 30 mm) Si applications a not as conformal as wafers, so we expect thick substrates could have gaps even in the absence of particles.  Most of our thick substrates are optically polished to much less than $\lambda = 632$ nm peak to valley over the 3 to 4 inch diameter clear aperture.

@@

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.95\columnwidth]{figs/VG03patternSchematic.pdf}}
\caption{Schematic of part VG03 \label{figVG03pattern}.  The 100 mm diameter VG03 is comprised of two 3.3 mm thick silicon pucks.  We patterned a 25 mm diameter hole on one of the pucks.  The depth was 4.0 $\mu$m as indicated by Dektak profilometry.  We also patterned a low-fill factor petal pattern surrounding the hole.  The purpose for these patterns was to have gaps of known dimensions against which we could test the optical metrology technique we describe in this paper.}
\end{figure}

We evaluated our metrology method by imbedding gaps of known sizes in the interface.  Here we describe the process of how we made the gaps and how we characterized their axial extent.

The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.  We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.  The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.  There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.

The three important characteristics of the gaps are their axial extent, their lateral areal extent, and their areal fill factor.  We bored holes in silicon with inductively coupled plasma etching.  We used two different gases with low and high etch rates.  CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 micron/minute.  The area exposed to the process gases was about 1 inch diameter circles, or 15 mm $\times$ 20 mm rectangles.  We protected the unexposed areas with thick photoresist.  We achieved etch depths of 15$\pm$5 nm and 4000$\pm$100 nm, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.  Table \ref{tabbondexper} has more information about the bonding samples.  Figure \ref{figVG03pattern} shows a schematic of VG03 and its pattern, as described in Table \ref{tabbondexper}.  In a second round of experiments we achieved gap sizes of 14$-$95 nm on four substrates with three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.  For measurements taken with the Veeco NT9100 Optical Profiler, uncertainties were constructed by inspecting the histogram of topology, as shown in Figure \ref{figVS20pattern}.  The large-scale distortions were removed from the topology by masking and flattening to ensure that the uncertainties accurately reflect the distribution of measured heights.  The stylus profiliometry measured values were simply assigned an uncertainty of 5\%, which was consistent with our intuition and experience.  Table \ref{tabbondexper} lists part numbers and descriptions for our bonded substrates.

@@

\begin{table}[h!]
\caption{UTexas Si bonding experiments \label{tabbondexper}}
\begin{center}
    \begin{tabular}{ c c c c c c c}
    \hline
    Name & Bonded/DSP & Diameter & Thickness & Pattern Type & Pattern Depth & Notes \\ 
    - & - & mm & mm & - & nm & - \\ 
        \hline
    VG01 & Bonded  & 100  &$1$ $\times$ 2 & None  & 2/13/2013 &  \\
    VG02 & Bonded & 100 & $3.3$ $\times$ 2 &  None  & 2/13/2013 &  \\
    VG03 & Bonded & 100 & $3.3$ $\times$ 2 &  Hole \&Petal & 4000$\pm$200 & See Fig. \ref{figVG03pattern}\\    
    VG04 & Bonded & 100 & $1$ $\times$ 2 &  Patchwork Islands & 15 & \\
    VG05 & DSP & 100 & 1&  - & - & -  \\
    VG06 & DSP & 100 & 3.5 &  - & - & - \\
    VG07 & - & - & - &  - & - & \\
    VG08 & - & - & - &  - & - & \\
    VG09 & Bonded & 75 & 1.5? & mesh C & 49 $\pm$6 & \\
    VG15 & Bonded & 100 & 1.5? &  mesh C, M, F  &14 $\pm$2 & \\
    VG17 & Bonded & 75 & 6.0? &  mesh C & 64 $\pm$6 nm & \\
    VS20 & Bonded & 100 & 1.5? &  mesh C, M, F & 95 $\pm$5 & \\
    \hline
    \end{tabular}
\end{center}
\end{table}

\section{Substrate preparation}
We cleaned the surfaces before bonding to minimize the interfacial particle density.  The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler.  We measured the large scale surface flatness of the pucks.  The Fisba2 interferometer has a 2 inch diameter beam at $\lambda=$632.8 nm.  We found a typical peak to valley surface flatness of 2 waves over the central 2 inch diameter.
